---
title: "work_with_data"
author: "Anatoliy"
date: "2023-09-03"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(dplyr)
library(tibble)
library(readr)
library(datasets)

#to include pivat_longer
install.packages("tidyverse")
library(tidyverse)

# to install flextable:
install.packages("flextable")
library(flextable)
```

## DATA

``` {r echo=FALSE}

data_tsv <- readr::read_tsv("data_tsv.tsv"
                        , na = c("", "NA")
                        , skip = 0
                        , n_max = Inf
                        , col_names = TRUE)
summary(data_tsv)
```

```{r}
mutated_data <- data_tsv %>%
  select(., `Группа`, `Рост`, where(is.numeric) & contains("E1")) %>%
  filter(., `Группа` == "Группа 1") %>%
  slice(., 1:10) %>%
  mutate(., `Рост` =  `Рост` / 10)
```

use tibble
```{r}
data_tibble <- as_tibble(data_tsv)

data_tibble$`Группа крови`

data_tibble[["Группа крови"]]
```

```{r}
#laizy <- tibble(var_first = 1:10, var_second = ifelse(var_first < 5, var_first + 100, var_first))
#laizy_2 <- tibble(`var 1` = 1:10, `var 2` = `var 1` * 100)
###laizy_3 <- tibble(var_1 = c(1:10) - 100, var_2 = 1:100)
###laizy_4 <- tibble(`var 2` = 10:1, `var 3` = `var 1` - 10)
###laizy_5 <- tibble(var_first = 1:10, var_first = ifelse(var_first < 5, var_first + 100, var_first))
###laizy_6 <- tibble(var = 1:10, var = var - 10000000)
```

### View()
to see tabels by eyes

```{r}
  View(data_tsv)
```

### tibble::add_column()

```{r}
#data %>% add_column(column_name = 1:10, .before = NULL, .after = NULL)
  data_tibble <- data_tibble %>%
    tibble::add_column(additional_column = 1, .before = 1)
```

### tibble::add_row()
```{r}
#data %>% add_row(var_1 = 1, var_2 = "value", .before = NULL, .after = NULL)
data_tibble <- data_tibble %>%
                  add_row(Группа = "Группа 1", `Возраст` = 110, .before = 1)
```

### dplyr::row_number()
Одна полезная функция, которая даёт возможность пронумеровать строки (добровольцев).
аналог "_n_" в sas

```{r}
data_tibble <- data_tibble %>%
                  mutate(ID = row_number())
```

### Concatination for tables

```{r}
#data_1 %>% bind_cols(data_2) %>% bind_cols(data_3)
data_1 <- tibble(var_1 = 1:10, var_2 = rep(c("Group 1", "Group_2"), 5))
data_2 <- tibble(var_2 = rbeta(1, 10, 5), var_3 = rnorm(10))
data_3 <- tibble(var_4 = 100:91, var_5 = rep(c("Yang", "Old"), 5))

concat_data <- data_1 %>% bind_cols(data_2) %>% bind_cols(data_3)
```

### Concatination for tables

```{r}
#data_1 %>% bind_rows(data_2) %>% bind_rows(data_3)
```

### JOIN

```{r}
data_1 <- tibble(var_1 = 1:8) %>% mutate(id = row_number())
data_2 <- tibble(var_2 = rnorm(10)) %>% mutate(`Subject ID` = row_number())

data_3 <- data_1 %>% left_join(data_2, by = c("id" = "Subject ID"))

```

### Group by

```{r}
#dplyr::group_by()

data_tibble %>%  group_by(Группа,  Пол)
```

### SPLIT

```{r}
#base::split()
data_tibble %>% split(~ `Группа`)
```

### Groupping by raws

```{r}
data_tibble %>% 
  rowwise() %>%
  mutate(`Cреднее_Баз_Два_Визита` = mean(c(`Базофилы_E1`, `Базофилы_E2`))) %>%
  ungroup() %>%
  select(`Базофилы_E1`, `Базофилы_E2`, `Cреднее_Баз_Два_Визита`)
```

```{r}
data_tsv %>% glimpse()
```

### Select
```{r}
#dplyr::select(), where()
data_tsv %>% select(`Группа крови`, `Рост`)
```

```{r}
# all bariables exept in list

data_tsv %>% select(!`Группа крови`)
```

```{r}
#tidyselect::contains(), tidyselect::matches(), tidyselect::starts_with(), tidyselect::ends_with()
data <- data_tsv
data %>% select(matches("_E\\d{1}"))
```


```{r}
# tidyselect::all_of(), tidyselect::any_of()
variables_data_1 <- c("Базофилы_E1", "Базофилы_E2")

variables_data_2 <- c("Базофилы_E1", "Базофилы_E2", "missing")

data %>% select(all_of(variables_data_1))
data %>% select(any_of(variables_data_2))
```

```{r}
#dplyr::slice()
data %>% slice(1:10)

data %>% slice_min(`Возраст`)

data %>% slice_max(`Возраст`)


```

```{r}

data %>% filter(`Возраст` <= median(`Возраст`))

data %>% filter(`Пол` == "Женский")

data %>% filter(`Группа крови` %in% c("A (II)", "B (III)") & `Группа` != "Группа 1")

data %>% filter(between(`Возраст`, 40, 100))

#soft filtration

data %>% filter(near(`Эозинофилы_E1`, 3.38, tol = 0.01))
```


#### Использование if_any() и if_all() для фильтрации сразу по нескольким переменным

```{r}
  data %>% filter(if_all(.cols = contains("Базофилы"), .fns = function(x) x > 1.5))

data %>% filter(if_any(.cols = contains("Базофилы"), .fns = function(x) x > 1.5))
```

```{r}
#dplyr::mutate()

data %>% mutate(`F_IV` = ifelse(`Пол` == "Женский" & `Группа крови` == "AB (IV)", "Y", "N")) %>%
  select(F_IV, everything()) %>%
  arrange(F_IV)
```

```{r}
#dplyr::case_when()
data %>% mutate(`Возрастная группа` = case_when(`Возраст` < 20  ~ "< 20",
                                                between(`Возраст`, 20, 30) ~ "20 - 30",
                                                `Возраст` > 30  ~ "> 30") %>% as.factor()) %>%
  select(`Возраст`, `Возрастная группа`)
```

```{r}

data %>% mutate(across(!contains("E1") & !c(`Группа`, `Возраст`, `Пол`, `Группа крови`), function(x) x ^ 2))
```

```{r}
# dplyr::summarise()/dplyr::summarize()
data %>%
  select(`Группа`, where(is.numeric))%>%
  group_by(`Группа`)%>%
  summarise(across(is.numeric, function(x) mean(x, na.rm = T)))

```

```{r}
#list of statistics
statistics <- list(
      `Количество субъектов` = ~length(.x),
      `Количество (есть данные)` = ~sum(!is.na(.x)),
      `Нет данных` = ~sum(is.na(.x)),
      `Ср. знач.` = ~ifelse(sum(!is.na(.x)) == 0, "Н/П*", mean(.x, na.rm = TRUE) %>% round(2) %>% as.character()),
      `Станд. отклон.` = ~ifelse(sum(!is.na(.x)) < 3, "Н/П*", sd(.x, na.rm = TRUE) %>% round(2) %>% as.character()),
      `95% ДИ для среднего` = ~sd(.x, na.rm = TRUE) %>% round(2) %>% as.character(),
      `мин. - макс.` = ~ifelse(sum(!is.na(.x)) == 0, "Н/П*", paste0(min(.x, na.rm = TRUE) %>% round(2), " - ", max(.x, na.rm = TRUE) %>% round(2))),
      `Медиана` = ~ifelse(sum(!is.na(.x)) == 0, "Н/П*", median(.x, na.rm = TRUE) %>% round(2) %>% as.character()),
      `Q1 - Q3` = ~ifelse(sum(!is.na(.x)) == 0, "Н/П*", paste0(quantile(.x, 0.25, na.rm = TRUE) %>% round(2), " - ", quantile(.x, 0.75, na.rm = TRUE) %>% round(2)))
)
```

```{r}
#create first tabel with descript stats

data %>%
  select(`Группа`, where(is.numeric))%>%
  group_by(`Группа`)%>%
  summarise(across(where(is.numeric), statistics))%>%
  mutate(across(everything(), as.character)) %>%
  pivot_longer(!`Группа`)%>%
  separate(name, into=c("Переменная", "Статистика"), sep="_", extra = "merge")%>%
  rename(`Значение`=value) %>%
```

```{r}
# Create second table with factor statistics

data %>% 
  select(`Группа`, where(is.character)) %>%
  mutate(`Группа крови` = `Группа крови` %>% replace_na("Нет данных") %>% as.factor()) %>%
  count(`Группа`, `Группа крови`) %>%
  group_by(`Группа`) %>%
  mutate(`PCTBYGR` = (n / sum(n)) %>%  round(4) %>% `*` (100) %>% str_c("%")) %>%
  ungroup() %>%
  mutate(`PCTBYPOP` = (n / sum(n)) %>%  round(4) %>% `*` (100) %>% str_c("%"))
```

```{r}
#flextable::flextable()
data %>%
  select(`Группа`, where(is.numeric))%>%
  group_by(`Группа`)%>%
  summarise(across(where(is.numeric), statistics))%>%
  mutate(across(everything(), as.character)) %>%
  pivot_longer(!`Группа`)%>%
  separate(name, into=c("Переменная", "Статистика"), sep="_", extra = "merge")%>%
  rename(`Значение`=value) %>%
  flextable()
```

```{r}
#flextable::theme_box(), flextable::theme_booktabs(), flextable::theme_alafoli()
data %>%
  select(`Группа`, where(is.numeric))%>%
  group_by(`Группа`)%>%
  summarise(across(where(is.numeric), statistics))%>%
  mutate(across(everything(), as.character)) %>%
  pivot_longer(!`Группа`)%>%
  separate(name, into=c("Переменная", "Статистика"), sep="_", extra = "merge")%>%
  rename(`Значение`=value) %>%
  flextable() %>%
  theme_box() %>%
  merge_v(c("Группа", "Переменная"))
```
# Add first Title for testing commit
# Add Second Title For Testing Commit
